### 2020-09-20 ecosystem-data-analysis.Rmd
ecosystem data

---
title: "WI FastPlants: Ecosystem in plant indicators"
author: "Claudia Solis-Lemus"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(car)
```

# Initial data: `ecosystem-data.csv`

### Reading the data

Reading in the data (had to remove manually the '#' sign from the column and save as csv file; also I removed the rows of "dead plant" and the text with the description of variables. All in the original file: `102314-Phenotypic-Plasticity-Data.xlsx`):
```{r}
dat = read.table("data/ecosystem-data.csv", sep=",",header=TRUE)
str(dat)
```

The data has 9 columns:
```{r}
head(dat)
```

Some of the variables that need to be numbers, as actually read as character due to some cell values being words.
```{r}
dat[dat$Flowers == "dropsy",]
dat2 = dat[dat$Flowers != "dropsy",]
dat2$Flowers = as.numeric(dat2$Flowers)
str(dat2)
```

## Data visualization

We will focus on visualizations for a dependent variable that is numeric (e.g. height, cot width, flowers, bracts) with respect to an independent variable that is also numeric (osmocote).

We will do the plots only for X=osmocote, Y=height, but these same plots can be done for Y=CotWidth or Y=Flowers or Y=Bracts, or any other numerical variable in future datasets.

```{r, echo=FALSE}
p2 = ggplot(dat2, aes(x = Osmocote, y = Height, fill = Osmocote)) +
  geom_point(pch = 21, alpha=0.3, position = position_jitterdodge(jitter.height=0.075, jitter.width=0.1))+
  theme(
        plot.title = element_text(hjust=0.5, size=rel(1.8)),
        axis.title.x = element_text(size=rel(1.8)),
        axis.title.y = element_text(size=rel(1.8), angle=90, vjust=0.5, hjust=0.5),
        axis.text.x = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
        axis.text.y = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
        panel.background = element_blank(),
        axis.line = element_line(colour = "grey")##,
        )
p2
```

We want to give people the option to add a linear regression line:
```{r, echo=FALSE}
p2 = ggplot(dat2, aes(x = Osmocote, y = Height, fill = Osmocote)) +
  geom_point(pch = 21, alpha=0.3, position = position_jitterdodge(jitter.height=0.075, jitter.width=0.1))+
  #ylim(c(1,6))+
  theme(
        plot.title = element_text(hjust=0.5, size=rel(1.8)),
        axis.title.x = element_text(size=rel(1.8)),
        axis.title.y = element_text(size=rel(1.8), angle=90, vjust=0.5, hjust=0.5),
        axis.text.x = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
        axis.text.y = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
        panel.background = element_blank(),
        axis.line = element_line(colour = "grey")##,
        )
p2
```