### 2020-10-08 ecosystem-data-analysis.Rmd

---
title: "WI FastPlants: Ecosystem in plant indicators"
author: "Claudia Solis-Lemus"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

# Initial data: `ecosystem-data.csv`

### Reading the data

Reading in the data (had to remove manually the '#' sign from the column and save as csv file; also I removed the rows of "dead plant" and the text with the description of variables. All in the original file: `102314-Phenotypic-Plasticity-Data.xlsx`):
```{r}
dat = read.table("data/ecosystem-data.csv", sep=",",header=TRUE)
str(dat)
```

The data has 9 columns:
```{r}
head(dat)
```

Some of the variables that need to be numbers, as actually read as character due to some cell values being words.
```{r}
dat[dat$Flowers == "dropsy",]
dat2 = dat[dat$Flowers != "dropsy",]
dat2$Flowers = as.numeric(dat2$Flowers)
str(dat2)
```

## Data visualization

We will focus on visualizations for a dependent variable that is numeric (e.g. height, cot width, flowers, bracts) with respect to an independent variable that is also numeric (osmocote).

We will do the plots only for X=osmocote, Y=height, but these same plots can be done for Y=CotWidth or Y=Flowers or Y=Bracts, or any other numerical variable in future datasets.

```{r, echo=FALSE}
p2 = ggplot(dat2, aes(x = Osmocote, y = Height, fill = Osmocote)) +
  geom_point(pch = 21, alpha=0.3, position = position_jitterdodge(jitter.height=0.075, jitter.width=0.1))+
  theme(
        plot.title = element_text(hjust=0.5, size=rel(1.8)),
        axis.title.x = element_text(size=rel(1.8)),
        axis.title.y = element_text(size=rel(1.8), angle=90, vjust=0.5, hjust=0.5),
        axis.text.x = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
        axis.text.y = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
        panel.background = element_blank(),
        axis.line = element_line(colour = "grey")##,
        )
p2
```

We want to give people the option to add a linear regression line:
```{r, echo=FALSE}
p2.2 = ggplot(dat2, aes(x = Osmocote, y = Height, fill = Osmocote)) +
  geom_point(pch = 21, alpha=0.3, position = position_jitterdodge(jitter.height=0.075, jitter.width=0.1))+
  geom_smooth(method='lm', formula= y~x)+
  theme(
        plot.title = element_text(hjust=0.5, size=rel(1.8)),
        axis.title.x = element_text(size=rel(1.8)),
        axis.title.y = element_text(size=rel(1.8), angle=90, vjust=0.5, hjust=0.5),
        axis.text.x = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
        axis.text.y = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
        panel.background = element_blank(),
        axis.line = element_line(colour = "grey")##,
        )
p2.2
```

We can also do violin plots if one of the variables in categorical. We do not have any independent categorical variable right now (we only have osmocotes), but we could pretend that we have a categorical variable by treating the values of osmocotes as categories:
```{r}
dat2$osmocote.cat = factor(dat2$Osmocote)
str(dat2)
```

```{r}
p4 = ggplot(dat2, aes(x=osmocote.cat, y=Height, fill=osmocote.cat))+geom_violin(alpha=0.5)+
  theme(
        plot.title = element_text(hjust=0.5, size=rel(1.8)),
        axis.title.x = element_text(size=rel(1.8)),
        axis.title.y = element_text(size=rel(1.8), angle=90, vjust=0.5, hjust=0.5),
        axis.text.x = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
        axis.text.y = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
        panel.background = element_blank(),
        axis.line = element_line(colour = "grey")##,
        )

p4
```